Fréchet Distance

### 2022-12-03 docs/paper.tex
refactor

%%% date
\date{November 27, 2022}

@@

\section*{Abstract}

\section{Introduction}
% motivation
At first, it seems that measuring the similarity between two curves
is a very theoretical problem, but it has many, not that obvious, practical applications.
It is used in Geo-Information-Systems (GIS) to 
quantify the similarity or quality of two maps, % cite
find patterns in vehicle traffic or animal tracking data 
by trajectory analysis, % cite
gesture recognition on touch devices like smartphones \cite{hu_research_2022}, % cite
and shape matching in the field of computer vision.

To quantify the similarity of two given curves (e.g. shapes, paths, or trajectories)
a distance metric can be used. 
We can't simply use a distance metric like euclidean distance 
because we don't want to compare points to points 
but instead compare curves to curves.

@@

% Fréchet distance motivation
In reality, the course of the curves is important for many applications. 
For example, in handwriting character recognition on digital input devices, especially in Chinese, % TODO: add more citations
gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
For this reason, the \textit{Fréchet distance}, named after the famous French mathematician René Maurice Fréchet, was introduced.

A common analogy for the Fréchet distance can be shown by a man walking his dog. 
Assume that both the man and the dog are moving along a predefined fixed curve. 
Both are free to adjust their speed as they wish, even to stop. 
However, they are not allowed to go backwards. 
We are looking for the minimum length of the dog leash 
that is necessary for both of them to reach their destination. 
The length of the dog leash is then exactly the Fréchet distance of the two curves.
A formal definition is given in \autoref{def_frechet_distance}.

@@

% TODO: Add example graphic of Frechet distance

Besides the previously mentioned applications, 
where the course of the curves is important, 
the Fréchet distance also finds applications in
(online) handwriting signature verification \cite{zheng_algorithm_2008,fang_research_2018},
map-matching tracking data for traffic control, routing and navigation \cite{wenk_addressing_2006,brakatsoulas_map-matching_2005},
and moving object analysis, e.g., detecting commuting patterns \cite{buchin_detecting_2011}.

@@

\subsection{Fréchet distance with speed limits}

\subsection{Fréchet Distance for Uncertain Curves}

@@

% Free Space Diagram
\begin{figure}[ht]
  \centering
  % TODO: update figure graphics
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
        \input{images/frechet_distance/line_segments.tex}
      }
      \caption{Line segments $P$ and $Q$}
      \label{fig_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
        \input{images/frechet_distance/free_space_single_cell.tex}
      }
      \caption{Free space diagram}
      \label{fig_free_space_diagram_example}
  \end{subfigure}
  \caption[Free space diagram example]{
    The free space diagram \autoref{fig_free_space_diagram_example} 
    for line segments $P$ and $Q$ and distance $\varepsilon$
    from \autoref{fig_free_space_lines_example} using euclidean distance.}
  \label{fig_free_space}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  % TODO: update figure graphics
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
        \input{images/frechet_distance/polygonal_curves.tex}
      }
      \caption{Polygonal curves $P$ and $Q$}
      \label{fig_complex_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
        \input{images/frechet_distance/free_space_polygonal_curves.tex}
      }
      \caption{Free space diagram}
      \label{fig_complex_free_space_diagram_example}
  \end{subfigure}
  \caption[Complex free space diagram example]{
    The free space diagram \autoref{fig_complex_free_space_diagram_example} 
    for polygonal curves $P$ and $Q$ from \autoref{fig_complex_free_space_lines_example} 
    and distance $\varepsilon = 0.73$ using euclidean distance.}
  \label{fig_complex_free_space}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  % TODO: Replace graphic with appropiate example
  \includegraphics[width=0.3\textwidth]{images/frechet_distance/square-ellipse-intersection.pdf}
  \caption[Legal intervals on free space cell boundary]{
    Legal intervals on the boundaries of a free space cell are shown with black white dashed lines. 
    There are at most 8 intersection points for an ellipse and a rectangle.}
  \label{fig_legal_boundary_intervals_of_a_cell}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  % TODO: Replace graphic with appropiate example
  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
  \caption{The monotonically increasing curve through the free space is computed by dynamic programming.}
  \label{fig_free_space_curve}
\end{figure}

@@

\subsection{Improved Version?}
\label{sec_improved_algorithm}
\cite{van_leusden_novel_2013}
\cite{buchin_four_2017}

@@

\section{Conclusion}
% Which version to use when

>>>>

%%% date
\date{December 05, 2022}

@@

% \section*{Abstract}

\section{Introduction}
% motivation
At first, it seems that measuring the similarity between two curves
is a very theoretical problem, but it has many, not that obvious, practical applications.
It is used in Geo-Information-Systems (GIS) to 
quantify the similarity or quality of two maps \cite{lyu_partial-frechet-distance-based_2022},
find patterns in vehicle traffic or animal tracking data 
by trajectory analysis \cite{buchin_detecting_2011},
gesture recognition on touch devices like smartphones \cite{hu_research_2022},
and shape matching in the field of computer vision.

To quantify the similarity of two given curves (e.g. shapes, paths, or trajectories)
a distance metric can be used. 
We can't simply use a distance metric like euclidean distance 
because we don't want to compare points to points 
but instead compare curves to curves. 
Therefore, we start by defining curves in \autoref{def_curve}.

@@

% Fréchet distance motivation
In reality, the course of the curves is important for many applications. 
For example, in handwriting character recognition on digital input devices, especially in Chinese \cite{chanin_david_hanzi_nodate},
gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
For this reason, the \textit{Fréchet distance}, named after the famous French mathematician René Maurice Fréchet, was introduced.

A common analogy for the Fréchet distance can be shown by a man walking his dog. 
Assume that both the man and the dog are moving along a predefined fixed curve. 
Both are free to adjust their speed as they wish, even to stop. 
However, they are not allowed to go backwards. 
We are looking for the minimum length of the dog leash 
that is necessary for both of them to reach their destination. 
The length of the dog leash is then exactly the Fréchet distance of the two curves.
A formal definition is given in \autoref{def_frechet_distance} 
along with a visual example in \autoref{fig_frechet_distance_example}.

@@

\begin{figure}[ht]
  \centering
  % TODO: Add example graphic of Frechet distance
  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
  \caption{The Fréchet distance is indicated by the black lines connecting the two polygonal curves.}
  \label{fig_frechet_distance_example}
\end{figure}

% TODO: merge with above mentions of applications?
Besides the previously mentioned applications, 
where the course of the curves is important, 
the Fréchet distance also finds applications in
(online) handwriting signature verification \cite{zheng_algorithm_2008,fang_research_2018},
map-matching tracking data for traffic control, routing and navigation \cite{wenk_addressing_2006,brakatsoulas_map-matching_2005},
and moving object analysis, e.g., detecting commuting patterns \cite{buchin_detecting_2011}.

@@

% \subsection{Fréchet distance with speed limits}

% \subsection{Fréchet Distance for Uncertain Curves}

@@

% Free Space Diagram
\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
        \input{images/frechet_distance/line_segments.tex}
      }
      \caption{Line segments $P$ and $Q$}
      \label{fig_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
        \input{images/frechet_distance/free_space_single_cell.tex}
      }
      \caption{Free space diagram}
      \label{fig_free_space_diagram_example}
  \end{subfigure}
  \caption[Free space diagram example]{
    The free space diagram \autoref{fig_free_space_diagram_example} 
    for line segments $P$ and $Q$ and distance $\varepsilon$
    from \autoref{fig_free_space_lines_example} using euclidean distance.}
  \label{fig_free_space}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
        \input{images/frechet_distance/polygonal_curves.tex}
      }
      \caption{Polygonal curves $P$ and $Q$}
      \label{fig_complex_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
        \input{images/frechet_distance/free_space_polygonal_curves.tex}
      }
      \caption{Free space diagram}
      \label{fig_complex_free_space_diagram_example}
  \end{subfigure}
  \caption[Complex free space diagram example]{
    The free space diagram \autoref{fig_complex_free_space_diagram_example} 
    for polygonal curves $P$ and $Q$ from \autoref{fig_complex_free_space_lines_example} 
    and distance $\varepsilon = 0.73$ using euclidean distance.}
  \label{fig_complex_free_space}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.3\textwidth]{images/frechet_distance/square-ellipse-intersection.pdf}
  \caption[Legal intervals on free space cell boundary]{
    Legal intervals on the boundaries of a free space cell are shown with black white dashed lines. 
    There are at most 8 intersection points for an ellipse and a rectangle.}
  \label{fig_legal_boundary_intervals_of_a_cell}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  % TODO: Replace graphic with appropiate example
  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
  \caption{The monotonically increasing curve (red) through the free space diagram.}
  \label{fig_free_space_curve}
\end{figure}

@@

% \subsection{Improved Version?}
% \label{sec_improved_algorithm}
% \cite{van_leusden_novel_2013}
% \cite{buchin_four_2017}

@@

\section{Conclusion}
% Which version to use when
This work gives an overview on the Fréchet distance, 
its variants and algorithms to compute them. 
Several differences between the variants have been described 
and evaluated. 
In conclusion, it can be stated that computing the continuous 
Fréchet distance is more computationally expensive compared 
to its discrete and weak variants.
However, computing the continuous variant can be done efficiently
in practice by binary searching the bits of the desired accuracy directly
instead of computing the formally correct value. 
Alternatively, the discrete Fréchet distance can be calculated, 
since it constrains the continuous distance 
from below and above. 
In this way, the tolerance range can be narrowed down.
